---
title: "2013_SPSSdata"
author: "Yunan Zhao"
date: "12/7/2021"
output: html_document
---

### Library
Here are the R packages required for processing the following codes
```{r setup, message=FALSE, warning=FALSE}
library(haven)
library(readr)
library(mice) # for missing value
library('yarrr') # for color palettes
library(tidyverse)
library(corrplot) # for correlation plot
library(table1) # for table generation
library(ggplot2)
library(ResourceSelection) # GOF for logistic regression model
library(LogisticDx) # GOF: Discrimination
```

### YRBSS Data
Studies of the YRBSS Data in 2013

Dependent Variable: whether the participant seriously consider attempting suicide? (binary variable)

```{r}
yrbs2013 <- read_sav("yrbs2013.sav")
#write.csv(yrbs2013, file="yrbs2013.csv",row.names = FALSE)
#yrbs2013 <- read_csv("yrbs2013.csv")
#View(yrbs2013)
```

Firstly, we create the subset for the demographics, including `age`, `sex`, `grade`, `race`, `height`, and `weight` of each participant.

In addition, we are also interested in the association between attempting suicide with other risk behaviors. Selecting variables fall into one of the following 5 categories:

1) Unintentional Injuries and Violence (e.g. driving while drinking alcohol, suicide behaviors)

2) Substance Use (e.g. smoking status, current alcohol use, marijuana use)

3) Sexual Behavior (e.g. sexual intercourse, condom usage)

4) Physical Inactivity (e.g. physical education class attendance)

5) Weight Management (e.g. Weight management intention, unhealthy weight control behaviors)

```{r}
id <- 1:dim(yrbs2013)[1]
yrbs2013_data <- cbind(id, yrbs2013[c(1:3,5:7,11,27,31,43,47,59,64,67:70,83)])
names(yrbs2013_data) <- c("id","age","sex","grade","race","height","weight",
                                  "drive_drink", "suicide",
                                  "smoke", "alcohol", "marijuana",
                                  "sexual","condom",
                                  "weight management","fast", "diet pills", "vomit/laxative",
                                  "PE"
                                  )
#write.csv(yrbs2013_data, file="yrbs2013_data.csv",row.names = FALSE)
```

### Data Wrangling
Though the YRBSS data is pre-processed by the YRBSS organization, to dig more into the data for statistical analysis and visualization, we will do data wrangling to make it in the format we could quickly make desired result.

```{r, warning=FALSE, message=FALSE}
yrbs2013_data <- read_csv("yrbs2013_data.csv")
#View(yrbs2013_data)
# Change "age" from group to exact age
yrbs2013_data$age2 <- yrbs2013_data$age
yrbs2013_data$age2[yrbs2013_data$age == 1] <- 12
yrbs2013_data$age2[yrbs2013_data$age == 2] <- 13
yrbs2013_data$age2[yrbs2013_data$age == 3] <- 14
yrbs2013_data$age2[yrbs2013_data$age == 4] <- 15
yrbs2013_data$age2[yrbs2013_data$age == 5] <- 16
yrbs2013_data$age2[yrbs2013_data$age == 6] <- 17
yrbs2013_data$age2[yrbs2013_data$age == 7] <- 18
# Change "sex" from number to character and mark each group
yrbs2013_data$sex2 <- yrbs2013_data$sex
yrbs2013_data$sex2[yrbs2013_data$sex == 1] <- "Female"
yrbs2013_data$sex2[yrbs2013_data$sex == 2] <- "Male"
# Change "race" from number to character and mark each group
yrbs2013_data$race2 <- rep("Multiple", dim(yrbs2013_data)[1])
yrbs2013_data$race2[yrbs2013_data$race == "A"] <- "Am Indian / Alaska Native"
yrbs2013_data$race2[yrbs2013_data$race == "B"] <- "Asian"
yrbs2013_data$race2[yrbs2013_data$race == "C"] <- "Black or African American"
yrbs2013_data$race2[yrbs2013_data$race == "D"] <- "Native Hawaiian/other PI"
yrbs2013_data$race2[yrbs2013_data$race == "E"] <- "White"
yrbs2013_data$race2[is.na(yrbs2013_data$race)] <- NA
# Calculate BMI as additional covariate
yrbs2013_data$BMI <- yrbs2013_data$weight/(yrbs2013_data$height)^2
# Change "drive_drink" from number to binary variable to indicate past driving with drinking experience
yrbs2013_data$drive_drink2 <- yrbs2013_data$drive_drink
yrbs2013_data$drive_drink2[yrbs2013_data$drive_drink == 1] <- 0
yrbs2013_data$drive_drink2[yrbs2013_data$drive_drink == 2 | yrbs2013_data$drive_drink == 3
                           |yrbs2013_data$drive_drink == 4 | yrbs2013_data$drive_drink == 5] <- 1
# Change "suicide" from number to character and mark each group
yrbs2013_data$suicide2 <- yrbs2013_data$suicide
yrbs2013_data$suicide2[yrbs2013_data$suicide == 2] <- 0
# Change "smoke" from number to character and mark each group
yrbs2013_data$smoke2 <- yrbs2013_data$smoke
yrbs2013_data$smoke2[yrbs2013_data$smoke == 2] <- 0
# Change "alcohol" from number to binary variable to indicate current alcohol usage
yrbs2013_data$alcohol2 <- yrbs2013_data$alcohol
yrbs2013_data$alcohol2[yrbs2013_data$alcohol == 1] <- 0
yrbs2013_data$alcohol2[yrbs2013_data$alcohol == 2|yrbs2013_data$alcohol == 3|yrbs2013_data$alcohol ==4|yrbs2013_data$alcohol ==5|yrbs2013_data$alcohol ==6|yrbs2013_data$alcohol ==7] <- 1
# Change "marijuana" from number to binary variable to indicate marijuana usage
yrbs2013_data$marijuana2 <- yrbs2013_data$marijuana
yrbs2013_data$marijuana2[yrbs2013_data$marijuana == 1] <- 0
yrbs2013_data$marijuana2[yrbs2013_data$marijuana == 2| yrbs2013_data$marijuana == 3|
                         yrbs2013_data$marijuana ==4| yrbs2013_data$marijuana ==5|
                         yrbs2013_data$marijuana ==6| yrbs2013_data$marijuana ==7] <- 1
# Change "sexual" from number to binary variable to indicate sexual behaviors
yrbs2013_data$sexual2 <- yrbs2013_data$sexual
yrbs2013_data$sexual2[yrbs2013_data$sexual == 2] <- 0
# Change "condom" from number to binary variable to indicate condom usage
yrbs2013_data$condom2 <- yrbs2013_data$condom
yrbs2013_data$condom2[yrbs2013_data$condom == 1 | yrbs2013_data$condom == 2] <- 0
yrbs2013_data$condom2[yrbs2013_data$condom == 3] <- 1
# Change "PE" from number to the number of days
yrbs2013_data$PE2 <- yrbs2013_data$PE - 1

# Change "weight management" from number to character and mark each group
yrbs2013_data$`weight management2` <- yrbs2013_data$`weight management`
yrbs2013_data$`weight management2`[yrbs2013_data$`weight management` == 1] <- "Lose weight"
yrbs2013_data$`weight management2`[yrbs2013_data$`weight management` == 2] <- "Gain weight"
yrbs2013_data$`weight management2`[yrbs2013_data$`weight management` == 3] <- "Stay the same weight"
yrbs2013_data$`weight management2`[yrbs2013_data$`weight management` == 4] <- "Not trying to do anything"
# Change "fast" from number to character and mark each group
yrbs2013_data$fast2<- yrbs2013_data$fast
yrbs2013_data$fast2[yrbs2013_data$fast == 1] <- "Yes"
yrbs2013_data$fast2[yrbs2013_data$fast == 2] <- "No"
# Change "diet pills" from number to character and mark each group
yrbs2013_data$`diet pills2`<- yrbs2013_data$`diet pills`
yrbs2013_data$`diet pills2`[yrbs2013_data$`diet pills` == 1] <- "Yes"
yrbs2013_data$`diet pills2`[yrbs2013_data$`diet pills` == 2] <- "No"
# Change "vomit/laxative" from number to character and mark each group
yrbs2013_data$`vomit/laxative2`<- yrbs2013_data$`vomit/laxative`
yrbs2013_data$`vomit/laxative2`[yrbs2013_data$`vomit/laxative` == 1] <- "Yes"
yrbs2013_data$`vomit/laxative2`[yrbs2013_data$`vomit/laxative` == 2] <- "No"
```


Then, we would generate `unhealthy weight control behavior` indicator to see whether the participant experienced any of the three "unhealthy weight control behavior"
```{r}
yrbs2013_data$UWCBs <- rep(0, dim(yrbs2013)[1])
yrbs2013_data$UWCBs[yrbs2013_data$fast == 1 | yrbs2013_data$`diet pills` == 1 | yrbs2013_data$`vomit/laxative` == 1] <- 1
yrbs2013_data$UWCBs[is.na(yrbs2013_data$fast) & is.na(yrbs2013_data$`diet pills`) & is.na(yrbs2013_data$`vomit/laxative`)] <- NA
write.csv(yrbs2013_data, file="yrbs2013_data2.csv",row.names = FALSE)

```

### Missing Values
Before examining data analysis, we would like to deal with the missing values in our data set (before data wrangling). 
```{r}
head(yrbs2013_data)
dim(yrbs2013_data)
sum(is.na(yrbs2013_data[1:24]))
```

Here, we print the number of missing values n for each column. There are 9147 total missing units within the data (before doing data wrangling), but for each column, the maximum of missing is 1198. Since this data has 13601 observations, and 1198 is only 8.81% of the whole data. Thus, we still have enough data to dealt with and no need to worry about the number of missing values.
```{r}
na_2013 <- md.pattern(yrbs2013_data[1:24], plot = FALSE, rotate.name = TRUE)
m_2013 <- na_2013[order(na_2013[,25], decreasing = TRUE),]
colnames(m_2013)[25] <- "sum"
m_2013[1,]
```
