---
output:
  html_document: default
  pdf_document: default
---
## Combining tables

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyverse)
library(ggrepel)
library(dslabs)
library(table1) # for table generation
library('yarrr') # for color palettes
library(mice) # for missing value
library(corrplot) # for correlation plot
library(ResourceSelection) # GOF for logistic regression model
library(LogisticDx) # GOF: Discrimination
library(caret)
```

The information we need for a given analysis may not be in just one table. Here we use a simple example to illustrate the general challenge of combining tables.

```{r}
data2013 <- read_csv("2013_NAomit.csv")
head(data2013)
mean(data2013$age2,na.rm=TRUE)
sd(data2013$age2,na.rm=TRUE)
```

Data Set Partition
```{r}
set.seed(1)
train.index <- sample(c(1:dim(data2013)[1]), dim(data2013)[1]*0.6)
valid.index <- setdiff(c(1:dim(data2013)[1]), train.index)
                      
train.df <- data2013[train.index, ]
valid.df <- data2013[valid.index, ]
#write.csv(train.df, file="train.df.csv",row.names = FALSE)
#write.csv(valid.df, file="valid.df.csv",row.names = FALSE)
```


#### Preliminary Statistics
```{r, warning=FALSE}
names(data2013)
data2013$suicide_chr[data2013$suicide2 == 1] <- 'Yes'
data2013$suicide_chr[data2013$suicide2 == 0] <- 'No'
table(data2013$suicide_chr)
round(mean(data2013$suicide2), 2)

data2013.fac <- data2013
data2013.fac$age2 <- as.factor(data2013.fac$age2)
data2013.fac$drive_drink2 <- as.factor(data2013.fac$drive_drink2)
data2013.fac$smoke2 <- as.factor(data2013.fac$smoke2)
data2013.fac$marijuana2 <- as.factor(data2013.fac$marijuana2)
data2013.fac$sexual2 <- as.factor(data2013.fac$sexual2)
data2013.fac$PE2 <- as.factor(data2013.fac$PE2)
data2013.fac$UWCBs <- as.factor(data2013.fac$UWCBs)


table1::label(data2013.fac$sex2) <- "Sex"
table1::label(data2013.fac$age2) <- "Age"
table1::units(data2013.fac$age2) <- "years"
table1::label(data2013.fac$race2) <- "Race"
table1::label(data2013.fac$drive_drink2) <- "Drive a car after drinking alcohol"
table1::label(data2013.fac$smoke2) <- "Smoke"
table1::label(data2013.fac$alcohol2) <- "Alcohol"
table1::label(data2013.fac$marijuana2) <- "Marijuana"
table1::label(data2013.fac$sexual2) <- "Sexual Behavior"
table1::label(data2013.fac$PE2) <- "Physical Education Attendance"
table1::units(data2013.fac$PE2) <- "#days/week"
table1::label(data2013.fac$weight.management2) <- "Weight Management"
table1::label(data2013.fac$UWCBs) <- "Unhealthy Weight Control Behaviors"

table1(~ sex2 + age2 + race2 + BMI + drive_drink2 + smoke2 + alcohol2 + marijuana2+sexual2+PE2+weight.management2 + UWCBs | suicide_chr, data=data2013.fac,
        caption = 'Table 2: Summary statistics of the dataset' )
```

```{r}
hist(data2013$age2,xlab='Age, years', main = "Histogram of Age")
hist(data2013$BMI,xlab='BMI, kg/m2', main = "Histogram of BMI")

```


#### Logistic Regression
From the step regression, the selected variables include race, drive_drink, sexual, weight management, PE, BMI, marijuana, smoke, alcohol, condom, age, sex, and UWCBs, variable `condom` is excluded.

```{r}
suicide.glm <- glm(suicide2 ~ sex2 + age2 + race2 + BMI +
                     drive_drink2 + smoke2 + alcohol2 + marijuana2 + sexual2 + condom2 + PE2 +
                     weight.management2 + UWCBs, data = train.df, family=binomial())
summary(suicide.glm)
train.complete <- na.omit(train.df)
suicide.glm.step = step(glm(suicide2 ~ 1,  data = train.complete, family=binomial),
                        ~ sex2 + age2 + race2 + BMI + drive_drink2 +
                          smoke2 + alcohol2 + marijuana2 + sexual2 + condom2 + PE2 + 
                          weight.management2 + UWCBs, direction = "both")
glm.point.est =  round(exp(coef(suicide.glm.step)),4)
glm.ci.est =  round(exp(confint(suicide.glm.step)),4)
glm.p.value = round(summary(suicide.glm.step)$coefficients[,4],4)
glm.est = cbind(glm.point.est,glm.ci.est,glm.p.value)

df = data.frame(glm.est)
colnames(df) = c("Odds Ratio", "2.5% CI", "97.5% CI", "p-value")
rownames(df)= c("Intercept","Unhealthy Weight Control Behaviors","Marijuana",
                "Sex","Smoke","Age",'Condom Usage',"Alcohol",'BMI',
                "Physical Education Attendance",'Weight management:Lose weight',
                'Weight management:Not trying to do anything',
                'Weight management:Stay the same level',
                "Drive a car after drinking alcohol",
                "Race:Asian",
                "Race:Black or African American",
                "Race:Multiple",
                "Race:Native Hawaiian/other PI",
                "Race:White")



df
```


```{r}
glm_train_prob <- predict(suicide.glm.step, train.df, type = "response")
glm_valid_prob <- predict(suicide.glm.step, valid.df, type = "response")

confusionMatrix(factor(ifelse(glm_train_prob>0.5,1,0)),factor(train.df$suicide2))
confusionMatrix(factor(ifelse(glm_valid_prob>0.5,1,0)),factor(valid.df$suicide2))
```


#### Naive Bayes Regressioon
```{r}
library(e1071)
suicide.nb <- naiveBayes(suicide2 ~ sex2 + age2 + race2 + BMI +
                     drive_drink2 + smoke2 + alcohol2 + marijuana2 + sexual2 + PE2 +
                     weight.management2 + UWCBs,
                         data = train.df)
nb_train_prob  <- predict(suicide.nb, newdata = train.df, type = "raw")
nb_train_class <- predict(suicide.nb, newdata = train.df)
nb_valid_prob  <- predict(suicide.nb, newdata = valid.df, type = "raw")
nb_valid_class <- predict(suicide.nb, newdata = valid.df)

confusionMatrix(factor(nb_train_class),factor(train.df$suicide2))
confusionMatrix(factor(nb_valid_class),factor(valid.df$suicide2))
```


```{r, warning=FALSE}
library(MASS)
suicide.lda <- lda(suicide2 ~ sex2 + age2 + race2 + BMI +
                     drive_drink2 + smoke2 + alcohol2 + marijuana2 + sexual2 + PE2 +
                     weight.management2 + UWCBs + condom2,
                         data = train.df)
lda_train_preds <- predict(suicide.lda, train.df)
confusionMatrix(data = as.factor(lda_train_preds$class),
                reference = as.factor(train.df$suicide2))
lda_valid_preds <- predict(suicide.lda, valid.df)
confusionMatrix(data = as.factor(lda_valid_preds$class),
                reference = as.factor(valid.df$suicide2))
```

```{r}
library(pROC)
suicide.lda <- lda(suicide2 ~ sex2 + age2 + race2 + BMI +
                     drive_drink2 + smoke2 + alcohol2 + marijuana2 + sexual2 + PE2 +
                     weight.management2 + UWCBs + condom2,
                         data = train.df)
lda_train_preds <- predict(suicide.lda, train.df)
confusionMatrix(data = as.factor(lda_train_preds$class),
                reference = as.factor(train.df$suicide2))
lda_valid_preds <- predict(suicide.lda, valid.df)
confusionMatrix(data = as.factor(lda_valid_preds$class),
                reference = as.factor(valid.df$suicide2))

roc_glm <- roc(valid.df$suicide2, nb_valid_prob[,2])
roc_lda <- roc(valid.df$suicide2, lda_valid_preds$posterior[,2])

plot(roc_glm)
plot(roc_lda, add=TRUE, col=2)

legend(0.2, 0.4, legend=c("GLM.stepwise", "LDA"),
       col=c("black", "red"), lty=1)
```

